---
title: "Class_10_markdown"
output: github_document
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

fna.data <- "data/WisconsinCancer.csv"
```{r}
fna.data <- "data/WisconsinCancer.csv"
wisc.df <- read.csv("WisconsinCancer.csv")


```

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```


Q1.. What type of object is returned from the read.csv() function? 
data frame

Q2. How many observations (i.e. patients) are in this dataset? 
569

Q3. How many of the observations have a malignant diagnosis? 
 212
 
Q4. How many variables/features in the data are suffixed with _mean? 
 there are 10

```{r}
nrow(wisc.data)
```
```{r}
table(wisc.df$diagnosis)
```

```{r}
colnames(wisc.df)
```

```{r}
grep("_mean",colnames(wisc.df),value = T)
```


```{r}
length(grep("_mean",colnames(wisc.df)))
```

## Enter Principal Component Analysis

First we need ot check whether our input dara should be scaled or not...

first we need to look at the mean value and standard deviation of the columns, if they are close then we do not need to use scale. If they are not...we need to scale!


```{r}
colMeans(wisc.data)
```


```{r}
round(apply(wisc.data,2,sd),2)
```


```{r}
wisc.pr <- prcomp(wisc.data,scale = T)
```

```{r}
summary(wisc.pr)
```

```{r}
biplot(wisc.pr)
```
```{r}
(wisc.pr)
```

"now that is one huge pile of crap"

```{r}
attributes(wisc.pr)
```

we want the $x component to make our pca plot

```{r}
plot(wisc.pr$x[,1:2],col=wisc.df$diagnosis)
```

We want the $x component to make our PCA plot!
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col=wisc.df$diagnosis)
```



```{r}
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.clust <- hclust((data.dist))
```

```{r}
wisc.pr.hc <- hclust(dist(wisc.pr$x[,1:3]),method = "ward.D2")
plot(wisc.pr.hc)
```


```{r}
grps <- cutree(wisc.pr.hc,k=2)
table(grps)
```


```{r}
table(grps,wisc.df$diagnosis)
```



```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2],col = grps)
```





```{r}
plot(wisc.clust)
abline(h=19,col = "red" ,lty=2)
```


wisc.hclust.clusters 
```{r}
wisc.hclust.clusters <- cutree((wisc.clust),k=4)
```






##Section 7

```{r}
new <- read.csv("new_samples.csv")
npc <- predict(wisc.pr, newdata=new)
npc
```


```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


