---
title: "Class_9_markdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-Clustering

```{r}
tmp <- c(rnorm(30,-3),rnorm(30,3))
x <- cbind((x=tmp), y=rev(tmp))

plot(x)
```

```{r}
km <- kmeans(x,2,20)
km
```

```{r}
attributes(km)
```


```{r}
plot(x,col=km$cluster)

points(km$centers,col = "blue", pch = 15, cex = 3)


```






## Hierarchical Clustering in R

the 'hclust()' function is the main Hierarchical clutering method in R and it **must** be passed a *distance* matrix as input, not your raw data!
```{r}
hc <- hclust(dist(x))

hc
```

```{r}
plot(hc)
```

```{r}
plot(hc)
abline(h = 6,col = "red")
abline(h=3.5, col="blue")
```


```{r}
cutree(hc,h=6)
```

```{r}
cutree(hc,h=3.5)
```
```{r}
table(cutree(hc,h=3.5))
```

```{r}
table(cutree(hc,k=5))
```





#Step 1. Generate some example data for clustering


```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd=0.3), ncol = 2), # c1
 matrix(rnorm(100, mean=1, sd=0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean=1, sd=0.3), # c3
 rnorm(50, mean=0, sd=0.3)), ncol = 2))
colnames(x) <- c("x", "y")

```


```{r}
plot(x)
```

```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

```

Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
Q. How does this compare to your known 'col' groups?

```{r}
hc_2 <- hclust(dist(x))
plot(hc_2)
```

```{r}
grps3 <- cutree(hc_2 , k = 3)
grps3
table(grps3)
```
```{r}
plot(x,col=grps3)

```


```{r}
grps2 <- cutree(hc_2 , k = 2)
grps2
```


```{r}
plot(x, col = grps2)
```

```{r}
table(grps3, col)
```

##Principal Component Analysis (PCA)
the main function in baase R for PCA is called 'prcomp()' here we will use PCA to examine the funny food that folks eat in the UK and N. Ireland

```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```

```{r}
dim(x)
```


```{r}
barplot(as.matrix(x),beside = F, col=rainbow(nrow(x)))
```


```{r}
pairs(x, col=rainbow(10), pch=16)
```


# PCA to the rescue

```{r}
pca <- prcomp(t(x))
pca
```
```{r}
summary(pca)
```


```{r}
plot(pca)
```

```{r}
attributes(pca)
```

```{r}
plot(pca$x[,1],pca$x[,2],xlab="PC1(67.4%)",ylab="PC2(29%)")
text(pca$x[,1],pca$x[,2], labels = colnames(x),col=c("black","red","blue","darkgreen"))
```

