---
title: "Class17"
author: "Alberto Carreno"
date: "3/4/2020"
output:
 github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here we analyze infection data for the 2019 novel Coronavirus COVID-19 (2019-nCoV) epidemic. The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus repository.

A CSV file is available here https://github.com/RamiKrispin/coronavirus-csv

```{r}
url <- "https://tinyurl.com/COVID-2019"
virus <- read.csv(url)

tail(virus)
```

>Q1. How many total infected cases are there around the world?

```{r}
total_cases <- sum(virus$cases)
total_cases

```

```{r}
table(virus$type)
```
```{r}
sum(virus$type == "death")
```

```{r}
inds <-  virus$type == 'death'
death_cases <- sum(virus[inds,"cases"])
death_cases
```


>Q2. how many deaths linked to infected cases have there been?

```{r}
inds <-  virus$type == 'death'
death_cases <- sum(virus[inds,"cases"])
death_cases
```

```{r}
round(death_cases/total_cases*100,2)
```

>Q3. What is the overall death rate?

```{r}
round(death_cases/total_cases*100,2)
```


```{r}
virus[,]
```
##Time for some HW!
```{r}
mainland <- virus[grep("Mainland China",virus$Country.Region),]
mainland

```
```{r}
mainland_2 <-  virus[virus$Country.Region == "Mainland China",]
mainland_2
```


> Q4. What is the death rate in "Mainland China"?

```{r}
mainland_total_cases <- sum(mainland$cases)
mainland_total_cases
```



```{r}
mainland_inds <-  mainland$type == 'death'
mainland_death_cases <- sum(mainland[mainland_inds,"cases"])
mainland_death_cases
```
```{r}
round(mainland_death_cases/mainland_total_cases,4)
```



> Q5. What is the death rate in Italy, Iran and the US?

```{r}
italy <-  virus[virus$Country.Region == "Italy",]
italy 
```
```{r}
Iran <-  virus[virus$Country.Region == "Iran",]
Iran 
```

```{r}
US <-  virus[virus$Country.Region == "US",]
US
```



```{r}
combined_total_cases <- sum(italy$cases,US$cases,Iran$cases)
combined_total_cases

```

```{r}
italy_inds <-  italy$type == 'death'
US_inds <-  US$type == 'death'
Iran_inds <-  Iran$type == 'death'
combined_death_cases <- sum(italy[italy_inds,"cases"],US[US_inds,"cases"],Iran[Iran_inds,"cases"])
combined_death_cases
```

```{r}
round(combined_death_cases/combined_total_cases,4)
```



```{r}
library(dplyr)
```


```{r}
cases_by_country <- virus %>% group_by(Country.Region) %>% 
  summarize(tot = sum(cases)) %>% 
  arrange(desc(tot))
head(cases_by_country)
```




