---
title: "class14"
author: "Alberto Carreno"
date: "2/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```
how many genes are we dealing with?

```{r}
nrow(counts)
```

```{r}
head(metadata)
```

we will compare the contrl vs treated count data contained in 'counts'. We will first extract the control columns by looking up their colnames in the 'metadata' object

```{r}
metadata
metadata$dex == "control"

```

```{r}
#sort for only the control condition by selecting the dex column, and specifying only rows that contain "control" written within the dex coluumn.
control <- metadata[metadata$dex == "control",]
control
```



```{r}
# access the counts coluns with conrtol$id 
counts[,control$id]
```



```{r}
#now find the mean value for each gene ie. each row
control.mean <- rowSums( counts[ ,control$id] )/ nrow(control)
names(control.mean) <- counts$ensgene
```



```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/nrow(control)
names(treated.mean) <- counts$ensgene
```

```{r}
mycounts <- data.frame(control.mean,treated.mean)
```
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```


```{r}
plot(meancounts[,1],meancounts[,2], xlab="Control", ylab="Treated")
```


```{r}
hist(mycounts$control.mean,breaks = 300)
```


```{r}
plot(mycounts,log="xy")
```

lets remove our zero count genes because we cant say anything about them from this dataset!
```{r}
x <- data.frame(c(1,3,10,0),
                c(1,3,0,0))
x
unique(which(x == 0, arr.ind = TRUE)[,"row"])
#which(x == 0)
```


ok find zero rows in our mycounts obect

```{r}
to.rm <- unique(which(mycounts == 0, arr.ind = TRUE)[,"row"])
newcounts <- mycounts[-to.rm,]
```

```{r}
nrow(newcounts)
```
```{r}
newcounts$log2fc <- log2(newcounts[,"treated.mean"]/newcounts[,"control.mean"])
head(newcounts)
```
A common rule of thumb in the field is to use a log2-fold change of greaterthan +2 as *upregulated* and less than -2 as *downregulated*

so, how many of our genes are upregulated upon this drug treatment

```{r}
sum(newcounts$log2fc > 2)
```
 how many are downregulated at this threshold?
 
```{r}
sum(newcounts$log2fc < -2)
```

##LETS USE DESeq2

```{r}
library(DESeq2)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
sizeFactors(dds)
```

```{r}
dispersions(dds)
```

```{r}
#results(dds)
```

```{r}
dds <- DESeq(dds)
```
```{r}
res <- results(dds)
res
```



```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
abline(v=c(-2,+2),lty = 2)
abline(h=log(0.05),lty = 2)
```

