---
title: "class_12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

>Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.
xray = 8.95 and EM = 2.72

```{r}
stats <-  read.csv("Data Export Summary.csv", row.names = 1)
stats
```

```{r}
total_sum <- sum(stats$Total)
```
```{r}
ans <- round(stats$Total / total_sum * 100,2)
names(ans) <- rownames(stats)
ans
```
>Q: What proportion of structures in the PDB are protein?

```{r}
round(sum(stats$Proteins)/total_sum *100,2)
```


>Q: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

there are 1289 hiv protease structures
## Section 2. Using bio3d to analyse protein structure data

Use the Bio3d package from CRAN


```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb

```

```{r}
pdb <- read.pdb("1AKE")
pdb
```
```{r}
aa <- get.seq("1ake_A")
```
```{r}
aa
```

#BLAST or hmmr search
```{r}
b <- blast.pdb(aa)
```
```{r}
hits <- plot(b)
```


```{r}
head(hits$pdb.id)
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```


```{r}
files <- get.pdb(hits$pdb.id, path = "pdbs" , split = TRUE , gzip = TRUE)
```


```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```



```{r}
library(bio3d.view)
library(rgl)

view(pdbs)
```

```{r}
anno <- pdb.annotate(ids)
unique(anno$source)
```

```{r}
#anno11
```

## Principal Component Analysis

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```


```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```



##GGPLOT TIME

```{r}
#Plotting results with ggplot2
library(ggplot2)
library(ggrepel)

df <- data.frame(x=pc.xray$z[,1], y=pc.xray$z[,2])
col <- as.factor(grps.rd)

p <- ggplot(df, aes(x, y)) +
 geom_point(aes(col=col), size=2) +
 xlab("PC1") +
 ylab("PC2") +
 scale_color_discrete(name="Clusters") +
 geom_text_repel(aes(label=ids))
p
```

```{r}
# NMA of all structures
modes <- nma(pdbs)

```
```{r}
plot(modes, pdbs, col=grps.rd)

```
